---
title: "New Mexico lizard weight NLS models"
author: "Kort Alexander"
date: "2/5/2022"
output: 
  html_document:
    code_folding: hide
    theme: spacelab
---

# Overview

This code will use non linear least squares to estimate parameters of a length to weight model for lizard populations in New Mexico.

Data citation: Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f

# Setup

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(broom)
library(kableExtra)
```

# Data input

```{r}
lizards <- read_csv(here("data", "lizard.csv"))
```

# Model function

```{r}
weight_model <- function(a, b, svl) {
  weight <- a * (svl)^b
  return(weight)
}
```

# Initial guesses calculation

```{r}
lizards <- lizards %>% 
  mutate(log_length = log(SV_length)) %>% 
  mutate(log_weight = log(weight))

guess_model <- lm(log_weight ~ log_length, data = lizards)
guess_model_table <- tidy(guess_model)

b_guess <- guess_model_table$estimate[2]
a_guess <- exp(guess_model_table$estimate[1]/b_guess)
```

# General NLS model creation and visualization

```{r}
lizards_nls <- nls(weight ~ weight_model(a, b, SV_length),
                   data = lizards,
                   start = list(a = a_guess, b = b_guess))

lizards_nls_table <- tidy(lizards_nls)

lizards <- lizards %>% 
  mutate(predicted_weight = weight_model(lizards_nls_table$estimate[1], lizards_nls_table$estimate[2], SV_length))

ggplot(lizards) +
  geom_point(aes(x = SV_length, y = predicted_weight, color = sex)) +
  labs(x = "Snout to vent length (mm)", y = "Predicted weight (g)", color = "Sex") +
  theme_minimal()
```

**Figure 1.** Weights of New Mexico lizards predicted by a general NLS model based on snout to vent length.

# General NLS model results

```{r}
lizards_nls_table <- lizards_nls_table %>% 
  mutate(p.value = ifelse(p.value < 0.001, "<0.001"))

colnames(lizards_nls_table) <- c("Term", "Estimate", "Standard error", "Statistic", "p value")
kable(lizards_nls_table, caption = "Table 1. General NLS model results.") %>% 
  kable_styling("striped")
```

# RMSE function

```{r}
rmse <- function(x, y) {
  rmse_result <- (x-y)^2 %>% mean() %>% sqrt()
  return(rmse_result)
}
```


# Western whiptail NLS model creation and visualization

```{r}
cnti_lizards <- lizards %>% 
  filter(spp == "CNTI", sex == "M")

cnti_nls <- nls(weight ~ weight_model(a, b, SV_length),
                data = cnti_lizards,
                start = list(a = a_guess, b = b_guess))

cnti_nls_table <- tidy(cnti_nls)

cnti_lizards <- cnti_lizards %>% 
  mutate(cnti_predicted_weight = weight_model(cnti_nls_table$estimate[1], cnti_nls_table$estimate[2], SV_length))

ggplot(cnti_lizards) +
  geom_point(aes(x = SV_length, y = predicted_weight, color = "General model prediction")) +
  geom_point(aes(x = SV_length, y = cnti_predicted_weight, color = "Specific model prediction")) +
  labs(x = "Snout to vent length (mm)", y = "Weight (g)", color = "") +
  theme_minimal()

cnti_lizards <- cnti_lizards %>% 
  mutate(rmse_general = rmse(predicted_weight, weight)) %>% 
  mutate(rmse_specific = rmse(cnti_predicted_weight, weight))

model_errors <- cnti_lizards %>% 
  summarize(general_mean_rmse = mean(rmse_general),
            specific_mean_rmse = mean(rmse_specific))
```

**Figure 2.** Comparison of predicted weights of male Western whiptail lizards between general NLS model and species specific NLS model. The general model has an average root-mean-square error of `r round(model_errors$general_mean_rmse, 2)`, while the species specific model has an average root-mean_square error of `r round(model_errors$specific_mean_rmse, 2)`. The Western whiptail specific model should be used because it has lower error on average.
